---
title: "assignment-5"
author: "Kelly Odion, Kirby Bartlett, Anna Calle"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and read in files
```{r, include = FALSE}
library(tidyverse)


faculty <- read_csv("faculty.csv")
grad_enroll <-read_csv("grad_enrollment.csv")
median_salary <- read_csv("median_salary.csv")
phd <- read_csv("phd_by_field.csv")


```



1) Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment
for males and females (including full-time/part-time and private/public universities) in the United
States from 1967 - 2015. Describe your results statistically, graphically and in text.
```{r}
# single linear regression?

# full time enrollment 

full_time <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Males_full_time)) +
  geom_line(aes(y = Females_full_time))
                      
full_time
#part time enrollment 


#private-public 



```

2) Shifts in female PhD recipients by field (1985, 2000, and 2015). Describe if and how there was
a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table,
and in text. Note: There are several ways that you can interpret this question. You are invited to decide
which you think is/are most interesting. Just be really clear about what you are asking/answering in
your report.
```{r}
#chi square? 
```

3) Male and female salaries for starting postdoctoral and other employment positions (2015).
Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:
Does median salary differ significantly between male and female starting postdoc positions? Does
median salary differ significantly between male and female PhD recipients in non-postdoc employment
positions?
```{r}

```

4) Exploring academic salaries for professors in U.S. colleges. Explore relationships between
variables in the ‘Faculty salary data (2008 - 2009 survey)’ dataset. Develop a model describing faculty
salary based on data for faculty sex, rank, years in current position, field, and number of years since
doctoral degree was earned. You should make decisions regarding which variables should remain in your
final model. Describe the results qualitatively and quantitatively (i.e., don’t just report the statistical
results of the model – make sure you describe interesting findings in text). You can also discuss any
concerns that you have with the model(s) you present, if any.
```{r}

```

