---
title: "assignment-5"
author: "Kelly Odion, Kirby Bartlett, Anna Calle"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and read in files
```{r, include = FALSE}
library(tidyverse)
library(stargazer)
library(MASS)
library(reshape2)

faculty <- read_csv("faculty.csv")
grad_enroll <-read_csv("grad_enrollment.csv")
median_salary <- read_csv("median_salary.csv")
phd <- read_csv("phd_by_field.csv")
phd_prop <-read_csv("phd_proportions.csv")


```



1) Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment
for males and females (including full-time/part-time and private/public universities) in the United
States from 1967 - 2015. Describe your results statistically, graphically and in text.

```{r}
#Data managment for part 1

#looking at total enrollment 
#model of all female students over time

grad_model_females <- lm(Total_Females ~ Year, data = grad_enroll)
grad_model_females

# equation: Enrollment = -58955502 + 30126(year) 

#explore model fit 

summary(grad_model_females)

correlation_females <- cor.test(grad_enroll$Year, grad_enroll$Total_Females)
correlation_females


#model of all male students over time

grad_model_males <- lm(Total_Males ~ Year, data = grad_enroll)
grad_model_males

#equation: Enrollment = -17112153 + 9069 (year)

#explore model fit
summary(grad_model_males)

correlation_males <- cor.test(grad_enroll$Year, grad_enroll$Total_Males)
correlation_males


#From 1967 to 2015, women enrollment in graduate programs rose by an average of 30,126 per year, while enrollment for men only rose by 9,069

#Summary of model stats for female model: the model signficantly predicts female graduate enrollment. p < 0.001. R squared is 0.982. Correlation is 0.9913, indicating a strong positive correlation. (b = 30126, t(47) = 51.66, p < 0.001), (F(1,47) = 2669, p < 0.001, R2 = 0.98)

#Summary of model stats for male model: the model signficantly predicts male graduate enrollment. p < 0.001. R squared is 0.8545. Correlation is 0.9568, indicating a strong positive correlation. (b = 9069, t(47) = 16.61, p < 0.001), (F(1,47) = 276, p < 0.001, R2 = 0.85) 

```

```{r}
#Graphs for part 1

#graph of model for females

grad_female_final_graph <- ggplot(grad_enroll, aes(x = Year, y = Total_Females)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
  theme_bw() +
  scale_x_continuous( expand = c(0,0), limits = c(1967, 2015)) +
   scale_y_continuous(expand = c(0,0), limits = c(0,1800000)) +
  labs(x = "Year", y = "US Annual Graduate School Enrollment- Women") +
  theme(text = element_text(family = "Times New Roman"))

grad_female_final_graph

# graph of model for males


grad_male_final_graph <- ggplot(grad_enroll, aes(x = Year, y = Total_Males)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
  theme_bw() +
  scale_x_continuous( expand = c(0,0), limits = c(1967, 2015)) +
   scale_y_continuous(expand = c(0,0), limits = c(0,1800000)) +
  labs(x = "Year", y = "US Annual Graduate School Enrollment- Men") +
   theme(text = element_text(family = "Times New Roman"))

grad_male_final_graph

```
__Figure 1. Total Female Enrollment in Graduate Programs (1967-2015).__ Annual enrollment by women in graduate programs between 1967 and 2015. Year significantly predicts graduate enrollment (b = 30126, t(47) = 51.66, p < 0.001) with a strong positive correlation (Pearson’s r = 0.99). The overall model signficantly explains women's enrollment each year (F(1,47) = 2669, p < 0.001, R2 = 0.98). Gray border indicates the 95% confidence interval for the mean predicted value. 

__Figure 2. Total Male Enrollment in Graduate Programs (1967-2015).__ Annual enrollment by men in graduate programs between 1967 and 2015. Year significantly predicts graduate enrollment (b = 9069, t(47) = 16.61, p < 0.001) with a strong positive correlation (Pearson’s r = 0.99). The overall model signficantly explains women's enrollment each year (F(1,47) = 276, p < 0.001, R2 = 0.85). Gray border indicates the 95% confidence interval for the mean predicted value.


1) continued: Extra graphs and info

```{r}
#non regression graphs in case we decide to use them

#total enrollment 
total_enroll <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Total_Males, color = "blue")) +
  geom_line(aes(y = Total_Females, color = "lightblue")) +
   scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1800000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))
                      
total_enroll
#full time enrollment 

full_time <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Males_full_time, color = "blue")) +
  geom_line(aes(y = Females_full_time, color = "lightblue")) +
  scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Full Time Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

full_time

#part time enrollment

part_time <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Males_part_time, color = "blue")) +
  geom_line(aes(y = Females_part_time, color = "lightblue")) +
  scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Part Time Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

part_time

#public vs private 

public_private <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Public, color = "blue")) +
  geom_line(aes(y = Private, color = "lightblue")) +
   scale_colour_discrete(name = "University Type", labels = c("Public", "Private")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,2000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

public_private
```

2) Shifts in female PhD recipients by field (1985, 2000, and 2015). Describe if and how there was
a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table,
and in text. Note: There are several ways that you can interpret this question. You are invited to decide
which you think is/are most interesting. Just be really clear about what you are asking/answering in
your report.
```{r}

#proportions of females in those four fields from 1985 to 2000 and from 2000 to 2015 

chisq <- chisq.test(phd_prop)
chisq

rownames(phd_prop) <- c(
"Physical sciences and earth sciences",
"Engineering",
"Education",
"Humanities and arts")


Department<-c(
"Physical sciences and earth sciences",
"Engineering",
"Education",
"Humanities and arts")
phd_names <- phd_prop %>% 
  cbind(phd_prop, Department)

library(corrplot)
corrplot(chisq$residuals, is.cor = FALSE)


```

Stacked bargraph for chisq
```{r}

prop_melt<-melt(phd_names)
prop_female <- ggplot(prop_melt, aes(x = variable, y = value)) +
  geom_col(aes(fill = Department))

prop_female 
```

3) Male and female salaries for starting postdoctoral and other employment positions (2015).
Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:
Does median salary differ significantly between male and female starting postdoc positions? Does
median salary differ significantly between male and female PhD recipients in non-postdoc employment
positions?
```{r}
#values are for the year 2015 

#employment positions 

#median femlae salary for starting employment 

med_female_employment <- median(median_salary$females_employment)
med_female_employment

#median male salary for starting employment 

med_male_employment <- median(median_salary$males_employment)
med_male_employment

# Compare male and female starting employment salaries using Mann Whitney 

test_employment <- wilcox.test(median_salary$males_employment, median_salary$females_employment, paired = TRUE)
test_employment


# Postdoc positions 

# median female salary for post doc positions 

med_female_postdoc <- median(median_salary$females_postdoc)
med_female_postdoc
# median male salary for post doc positions 

med_male_postdoc <- median(median_salary$males_postdoc)
med_male_postdoc
# Compare male and female startig postdoc salaries using Mann Whitley

test_postdoc <- wilcox.test(median_salary$males_postdoc, median_salary$females_postdoc, paired = TRUE)

test_postdoc

#create boxplot 

boxplot_data <- melt(median_salary)

employment_finalized <- ggplot(boxplot_data, aes(x = variable, y = value)) +
  geom_boxplot() +
  labs(y = "Annual Salary ($)", x = "") +
  geom_jitter(width = 0.1, aes(color = variable)) +
  scale_color_manual(values = c("darkblue","lightblue", "darkred", "red"), 
                     labels=c("Employed Males","Employed Females","Male Postdocs","Female Postdocs"), 
                     name = "Group") +
  scale_x_discrete(labels=c("Employed Males","Employed Females","Male Postdocs","Female Postdocs")) +
  theme_classic() +
   theme(text = element_text(family = "Times New Roman"))

employment_finalized

#Based on our results, there was not a significant difference in male and female salaires in 2015 in postdoc appointments, but there was in traditional employment settings (V = 101, p = 0.0025). Actual difference in median salaries for employment positions was $3,417. Actual difference in median salaries for post doc positions was $3,000


```
__Figure 2. Salaries for Female and Male doctorate recipients (2015).__ Annual salaries for men and women in traditional employment positions and in postdoctorate positions. bars indictae median values, boxes indiciates interquartile range. Data points indiciate individual observations. 


4) Exploring academic salaries for professors in U.S. colleges. Explore relationships between
variables in the ‘Faculty salary data (2008 - 2009 survey)’ dataset. Develop a model describing faculty
salary based on data for faculty sex, rank, years in current position, field, and number of years since
doctoral degree was earned. You should make decisions regarding which variables should remain in your
final model. Describe the results qualitatively and quantitatively (i.e., don’t just report the statistical
results of the model – make sure you describe interesting findings in text). You can also discuss any
concerns that you have with the model(s) you present, if any.
```{r}
# Use multiple linear regression 
 
# Saturated model
professor_lm1 <- lm(salary ~ faculty_rank + discipline + years_since_phd + years_faculty_service + sex, data = faculty)

summary(professor_lm1)

```
```{r}
professor_stepwise <- stepAIC(professor_lm1, direction = "both", 
                      trace = FALSE)
summary(professor_stepwise)
```

```{r}
# Updated model
# Excluded years since PhD because its largely expalined by years of faculty service

professor_lm2 <- lm(salary ~ faculty_rank + discipline + years_since_phd + sex, data = faculty)

summary(professor_lm2)

# Model diagnostics

plot(professor_lm2)

# Heteroscedasticity and residuals normality look OK

#Residual standard error: 22660 on 391 degrees of freedom
#Multiple R-squared:  0.4472,	Adjusted R-squared:  0.4401 
#F-statistic: 63.27 on 5 and 391 DF,  p-value: < 2.2e-16

```

```{r}
# AIC values
sat_aic <- AIC(professor_lm1)
final_aic <- AIC(professor_lm2)

```

```{r}
# Regression table
lm_table <- stargazer(professor_lm1, professor_lm2, type = "html")
lm_table
```

