---
title: "assignment-5"
author: "Kelly Odion, Kirby Bartlett, Anna Calle"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and read in files
```{r, include = FALSE}
library(tidyverse)


faculty <- read_csv("faculty.csv")
grad_enroll <-read_csv("grad_enrollment.csv")
median_salary <- read_csv("median_salary.csv")
phd <- read_csv("phd_by_field.csv")
phd_prop <-read_csv("phd_proportions.csv")


```



1) Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment
for males and females (including full-time/part-time and private/public universities) in the United
States from 1967 - 2015. Describe your results statistically, graphically and in text.
```{r}

#total enrollment 

total_enroll <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Total_Males, color = "blue")) +
  geom_line(aes(y = Total_Females, color = "lightblue")) +
   scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1800000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))
                      
total_enroll
#full time enrollment 

full_time <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Males_full_time, color = "blue")) +
  geom_line(aes(y = Females_full_time, color = "lightblue")) +
  scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Full Time Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

full_time

#part time enrollment

part_time <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Males_part_time, color = "blue")) +
  geom_line(aes(y = Females_part_time, color = "lightblue")) +
   scale_colour_discrete(name = "Student Type", labels = c("Males", "Females")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Part Time Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,1000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

part_time

#public vs private 

public_private <- ggplot(grad_enroll, aes(x = Year)) +
  geom_line(aes(y = Public, color = "blue")) +
  geom_line(aes(y = Private, color = "lightblue")) +
   scale_colour_discrete(name = "University Type", labels = c("Public", "Private")) +
  theme_classic() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(y = "Total Enrollment", x= "Year") +
   theme(axis.text=element_text(size=11))+
  theme(axis.title=element_text(size=12))+
  scale_y_continuous(expand = c(0,0), limits = c(0,2000000))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))

public_private

```

2) Shifts in female PhD recipients by field (1985, 2000, and 2015). Describe if and how there was
a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table,
and in text. Note: There are several ways that you can interpret this question. You are invited to decide
which you think is/are most interesting. Just be really clear about what you are asking/answering in
your report.
```{r}
#chi square? 
#proportions of females in those four fields from 1985 to 2000 and from 2000 to 2015 



```

3) Male and female salaries for starting postdoctoral and other employment positions (2015).
Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:
Does median salary differ significantly between male and female starting postdoc positions? Does
median salary differ significantly between male and female PhD recipients in non-postdoc employment
positions?
```{r}
#values are for the year 2015 

#employment positions 

#median femlae salary for starting employment 

med_female_employment <- median(median_salary$females_employment)
med_female_employment

#median male salary for starting employment 

med_male_employment <- median(median_salary$males_employment)
med_male_employment

# Compare male and female starting employment salaries using Mann Whitney 

test_employment <- wilcox.test(median_salary$males_employment, median_salary$females_employment)
test_employment


# Postdoc positions 

# median female salary for post doc positions 

med_female_postdoc <- median(median_salary$females_postdoc)
med_female_postdoc
# median male salary for post doc positions 

med_male_postdoc <- median(median_salary$males_postdoc)
med_male_postdoc
# Compare male and female startig postdoc salaries using Mann Whitley

test_postdoc <- wilcox.test(median_salary$males_postdoc, median_salary$females_postdoc)

test_postdoc


#Based on our results, there was not a significant difference in male and female salaires in 2015 either in postdoc appointments or in traditional employment settings (W = 117, p = 0.8671). Actual difference in median salaries for employment positions was $3,417. Actual difference in median salaries for post doc positions was $3,000


```

4) Exploring academic salaries for professors in U.S. colleges. Explore relationships between
variables in the ‘Faculty salary data (2008 - 2009 survey)’ dataset. Develop a model describing faculty
salary based on data for faculty sex, rank, years in current position, field, and number of years since
doctoral degree was earned. You should make decisions regarding which variables should remain in your
final model. Describe the results qualitatively and quantitatively (i.e., don’t just report the statistical
results of the model – make sure you describe interesting findings in text). You can also discuss any
concerns that you have with the model(s) you present, if any.
```{r}
# use multiple linear regression 
 
professor_lm1 <- lm(salary ~ faculty_rank, discipline, years_since_phd, years_faculty_service, sex, data = faculty)


```

